Depression and suicidality as costly signals of need

### 2021-04-06 Paper.rmd

The need to explain suicide's connection to depression is often overlooked by explanations for major depression, particularly those which propose non-signaling functions. One explanation for their connection, is that both suicidality and depression may function to limit negative fitness outcomes in times of substantial adversity [@andrews2006; @Hagen2002; @Hagen2003; @HagenEtAl2008; @nock2008; @Rosenthal1993; @Stengel1956; @SymeEtAl2016], with costlier signals helping elicit support when the payoffs to faking need are high [@Rosenthal1993]. For this hypothesis to be correct, instrumental support would need to be provided frequently enough with large enough impacts so that the costs of signaling are outweighed by the benefits on average. Although possible benefits to depressive states are rarely examined, we found exposure to depression and suicide attempts resulted in greater belief in need and increased likelihood of support in fictional scenarios. This coupled with findings that depression can be simultaneously aversive and elicit support [@HokansonEtAl1986; @stephens1987], suggests that more attention should be paid to the responses of others to depressed individuals, especially from those the depressed are in conflict with and those with the capability to provide meaningful benefits.

>>>>

The need to explain suicide's connection to depression is often overlooked by explanations for major depression, particularly those which propose non-signaling functions. One explanation for their connection, is that both suicidality and depression may function to limit negative fitness outcomes in times of substantial adversity [@andrews2006; @Hagen2002; @Hagen2003; @HagenEtAl2008; @nock2008; @Rosenthal1993; @Stengel1956; @SymeEtAl2016], with costlier signals being required to elicit support when the payoffs to faking need are high enough so that less costly signals are not reliable indicators of one's need [@Rosenthal1993]. For this hypothesis to be correct, instrumental support would need to be provided frequently enough with large enough impacts so that the costs of signaling are outweighed by the benefits on average. Although possible benefits to depressive states are rarely examined, we found exposure to depression and suicide attempts resulted in greater belief in need and increased likelihood of support in fictional scenarios. This coupled with findings that depression can be simultaneously aversive and elicit support [@HokansonEtAl1986; @stephens1987], suggests that more attention should be paid to the responses of others to depressed individuals, especially from those the depressed are in conflict with and those with the capability to provide meaningful benefits.
